Detection of an Intrinsic Color Parameter within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2016-09-30 forgerard.tex
commit

\subsection{Results}
\label{results:sec}
The posterior of the model parameters is evaluated using Hamiltonian Monte Carlo as implemented in
STAN \citep{stan}. STAN provides output statistics to assess
the convergence of the output Markov chains.
\textcolor{red}{The 
potential scale reduction statistic, $\hat{R}$ \citep{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multilple sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{eff}$ is an estimate of the number of independent draws.
The output gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{eff} \gg 100$ for all parameters, indicating that are all densely sampled.}
Empirically, the confidence regions are localized and unimodal as is seen in  Figures~\ref{global1:fig}, \ref{global2:fig}, \ref{global3:fig}, \ref{global4:fig},
\ref{global5:fig}.  We find no evidence that
the Monte Carlo chains have not converged to the stationary posterior distribution.
We have rurun the analysis with a variety of initial conditions, including one with all the $\gamma$'s and $\delta$'s equal to zero, except for a small positive 
$\gamma_0$ and small negative $\delta_0$ to satisfy our limit conditions.  Parameter credible intervals
remain consistent within uncertainty.

@@

For each of the five filters, the 68\%  equal-tailed credible intervals for the global parameters $\alpha$, $\beta$, $\eta$, $R$, $R_\delta$, and $\sigma$
(and other interesting derived parameters)
are given in Table~\ref{global:tab}.
\color{red}
As the $\gamma$ parameters are free up to a multiplicative constant and the highly non-zero values of $\gamma^i_V$,
their results are also shown in terms of $\gamma^i_X/\gamma^i_V-1$.
\color{black}
Contours of the posterior surface for parameter pairs grouped by filter are shown in Figures~\ref{global1:fig} -- \ref{global5:fig} .

@@

The extrinsic parameters $\gamma$ attributed to dust are fit for without reference to a specific dust model:
we now do so in a fit to the 
model of \citet{1999PASP..111...63F},
taking the fixed 3693,  4369,  5287,  6319, 7611 \AA\  for the effective wavelengths of the supernova flux through the bands.
Figure~\ref{fitz:fig} shows the   68\% credible intervals of $\gamma$ and our result
$R_V=2.97^{+0.18}_{-0.17}$..  
We do the analysis in $\gamma$, as opposed to $R_V$ say, because the posteriors of $\gamma$ are well-approximated by a normal distribution that is used for the fit.
The posterior distributions for $R_{\gamma X}=\frac{\gamma_X}{\gamma_B-\gamma_V}$ are shown in Figure~\ref{rx:fig}.
\color{black}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=3.2in]{output11/fitz.pdf}
   \caption{\textcolor{red}{68\% credible intervals of $\gamma_X$ and the best-fit to the  \citet{1999PASP..111...63F}  model $R^F_V=2.97$.}
   \label{fitz:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/rx_corner.pdf}
      \includegraphics[width=2.8in]{output11/rxdelta_corner.pdf} 
   \caption{Posterior contours for Left:  $R_X=\frac{\gamma_X}{\gamma_B-\gamma_V}$;  Right:  $R_{\delta X}=\frac{\delta_X}{\delta_B-\delta_V}$.
   \label{rx:fig}}
\end{figure}

@@

None of the 20000 links of 
our Monte Carlo chains for $\delta$ exceed 0 (see Figures~\ref{global1:fig}--\ref{global5:fig}), we thus claim detection of the
intrinsic parameter on supernova magnitudes
with probability $(1-5\times 10^{-5})$.
\textcolor{red}{
The values of $\delta_X/\delta_V$, plotted in  Figure~\ref{deltaratio:fig}, are not monotonic with color: they increase from the $U$ to $V$, and then decrease
from $V$ to $I$,
a consequence
of which is that
a reddening in $B-V$ comes with a blueing in $V-R$.
Also apparent is that a reddening of $B-V$ is associated with a brightening of the supernova.
The posterior distributions for $R_{\delta X}=\frac{\delta_X}{\delta_B-\delta_V}$ are shown in Figure~\ref{rx:fig}.
Although the posteriors of $\delta$ are close to Gaussian, $\delta_B-\delta_V$ can approach zero, resulting in the
elongated tail in negative  $R_{\delta X}$.
All values are significantly negative.}
These are not the behaviors expected from normal dust,
circumstellar dust suggested as a source of SN extinction \citep{2005ApJ...635L..33W,2008ApJ...686L.103G,
2015ApJ...807L..26G}, or a blackbody in the Raleigh-Jeans tail, and so may be indicative of
more complicated underlying supernova physics.

@@

Non-trivial residual magnitude dispersions are captured in $C_c$.   Figure~\ref{sigma:fig} shows the confidence regions for $\sigma$, the
square root of the diagonal elements of $C_c$.  The residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
The off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix.
Assembling
a new set of Cholesky factors based on the distributions of each individual factor will not generally satisfy the condition of representing a correlation matrix.  
To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
chain, element by element.
For $UBVRI$ the matrix is
\begin{equation}
\begin{pmatrix}
0.0040 & 0.0011 & -0.0002 & -0.0000 & 0.0003 \cr
0.0011 & 0.0012 & 0.0002 & -0.0000 & -0.0006 \cr
-0.0002 & 0.0002 & 0.0004 & 0.0001 & -0.0002 \cr
-0.0000 & -0.0000 & 0.0001 & 0.0002 & 0.0002 \cr
0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
 \end{pmatrix} \text{mag}^2.
 \label{mag_cov:eqn}
 \end{equation}
The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
expressed as the standard deviations and
 correlation matrix
 \begin{equation}
 \begin{pmatrix}
0.069 , 0.035 , 0.023 , 0.053
  \end{pmatrix} \text{mag}
 \label{color_sd:eqn}
   \end{equation}
 \begin{equation}
\begin{pmatrix}
1.000 & 0.624 & -0.318 & -0.287 \cr
0.624 & 1.000 & -0.180 & 0.104 \cr
-0.318 & -0.180 & 1.000 & 0.629 \cr
-0.287 & 0.104 & 0.629 & 1.000
  \end{pmatrix}.
  \label{color_cor:eqn}
 \end{equation}

 \begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output11/sigma_corner.pdf} 
   \caption{Posterior contours for the parameters $\sigma$, the square root of the diagonal elements of $C_c$.
   \label{sigma:fig}}
\end{figure}

Each supernova is described by its parameters $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_\gamma(B-V)$, and $E_\delta(B-V)$, as well as its grey offset
$\Delta$: their distributions for all Monte Carlo links for all supernovae are shown in Figure~\ref{perobject:fig}.
There is a core concentration in the  parameter-space, with around ten objects that occupy its outskirts.
Many outliers appear in the red tail of $E_\gamma(B-V)$, as would be expected for the infrequent selection of supernovae
heavily extinguished by host-galaxy dust.
Outliers  are also clearly distinguished in  $EW_{Ca}$--$\lambda_{Si}$ space.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output11/perobject_corner.pdf} 
   \caption{Distributions for the supernova parameters $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_\gamma(B-V)$, and $E_\delta(B-V)$, as well as the grey offset
$\Delta$.  All links are plotted, so that each supernova contributes a cloud of points.
   \label{perobject:fig}}
\end{figure}

The Pearson correlation coefficients for $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_\gamma(B-V)$, and $E_\delta(B-V)$ are given in the matrix
\begin{equation}
\begin{pmatrix}
1.000 & 0.001 & -0.059 & -0.045 & 0.105 & 0.007 \cr
0.001 & 1.000 & 0.115 & -0.252 & -0.066 & -0.003 \cr
-0.059 & 0.115 & 1.000 & -0.140 & -0.163 & -0.024 \cr
-0.045 & -0.252 & -0.140 & 1.000 & 0.040 & 0.027 \cr
0.105 & -0.066 & -0.163 & 0.040 & 1.000 & 0.055 \cr
0.007 & -0.003 & -0.024 & 0.027 & 0.055 & 1.000
\end{pmatrix}.
\end{equation}
\color{red}
We find weak correlations between our color-excess parameters and the input features, the strongest being $-0.163$ between
$E_\gamma(B-V)$ and $EW_{Si}$;
\color{black}
recall that $EW_{Si}$ is correlated with light-curve shape, which is correlated with host-galaxy (including dust) properties 
\citep{2003MNRAS.340.1057S}.
Although $E_\gamma(B-V)$ and $E_\delta(B-V)$ are determined simultaneously for each supernova from colors, they have a low correlation coefficient of 0.062;
this low correlation, and the broad distribution in $E_\delta(B-V)$ for low $E_\gamma(B-V)$,  indicate that what we have been calling intrinsic
is not related to extrinsic dust.
An extensive discussion of the correlations between the spectral features of the SNfactory data set can be found in \citet{chotard:thesis}
and \citet{leget:thesis}.

@@

\section{Interpreting $k$ and $D$ As Parameters of a Single Dust Model}
%ipdb> print y, y-ymin, ymax-y
%2.23720862938 0.00612061774367 0.00623996846008
%ipdb> (y,ymin,ymax)  = numpy.percentile(samples[:,1],(50,50-34,50+34)) 
%ipdb> print y, y-ymin, ymax-y
%-3.46707773157e-05 0.000555359008477 0.000616592111953
Until now we have considered $k$ and $D$ as independent external and internal parameters.  However, they may be both
parameters of a single underlying dust.  Indeed the CCM dust law can be exactly written as a two-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V).
\end{equation}
The Fitzpatrick model can also approximately be described by a linear model in the range of $R_V$ and wavelength under consideration:
the results shown here vary only at the percent level when making this approximation.
Similarly our model can be expressed as
\begin{equation}
A_X = \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2}  E_{\gamma^0}(B-V) +  \frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2}  E_{\gamma^1}(B-V).
\end{equation}

>>>>

\subsection{Results}
\label{results:sec}
The posterior of the model parameters is evaluated using Hamiltonian Monte Carlo as implemented in
STAN \citep{stan}. STAN provides output statistics to assess
the convergence of the output Markov chains.
\textcolor{red}{The 
potential scale reduction statistic, $\hat{R}$ \citep{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multilple sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{eff}$ is an estimate of the number of independent draws.
The output gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{eff} \gg 100$ for all parameters, indicating that are all densely sampled.}
Empirically, the confidence regions are localized and unimodal as is seen in  Figures~\ref{global1:fig}, \ref{global2:fig}, \ref{global3:fig}, \ref{global4:fig},
\ref{global5:fig}.  We find no evidence that
the Monte Carlo chains have not converged to the stationary posterior distribution.
We have rurun the analysis with a variety of initial conditions, including one with all the $\gamma$'s equal to zero, except for a small positive 
$\gamma^0_0$ and small negative $\gamma^1_0$ to satisfy our limit conditions.  Parameter credible intervals
remain consistent within uncertainty.

@@

For each of the five filters, the 68\%  equal-tailed credible intervals for the global parameters $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$
are given in Table~\ref{global:tab}.
\color{red}
As the $\gamma$ parameters are free up to a multiplicative constant and have non-zero values of $\gamma^i_V$,
their results are also shown in terms of $\gamma^i_X/\gamma^i_V-1$.
\color{black}
Contours of the posterior surface for parameter pairs grouped by filter are shown in Figures~\ref{global1:fig} -- \ref{global5:fig} .

@@

Non-trivial residual magnitude dispersions are captured in $C_c$.
% Figure~\ref{sigma:fig} shows the confidence regions for $\sigma$, the square root of the diagonal elements of $C_c$. 
The residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
The off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix.
Assembling
a new set of Cholesky factors based on the distributions of each individual factor will not generally satisfy the condition of representing a correlation matrix.  
To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
chain, element by element.
For $UBVRI$ the matrix is
\begin{equation}
\begin{pmatrix}
0.0038 & 0.0010 & -0.0002 & -0.0000 & 0.0003 \\
0.0010 & 0.0011 & 0.0002 & -0.0000 & -0.0006 \\
-0.0002 & 0.0002 & 0.0004 & 0.0000 & -0.0002 \\
-0.0000 & -0.0000 & 0.0000 & 0.0002 & 0.0002 \\
0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
 \end{pmatrix} \text{mag}^2.
 \label{mag_cov:eqn}
 \end{equation}
The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
expressed as the standard deviations and
 correlation matrix
 \begin{equation}
 \begin{pmatrix}
0.068 , 0.034 , 0.023 , 0.053
  \end{pmatrix} \text{mag}
 \label{color_sd:eqn}
   \end{equation}
 \begin{equation}
\begin{pmatrix}
1.000 & 0.615 & -0.334 & -0.296 \\
0.615 & 1.000 & -0.193 & 0.099 \\
-0.334 & -0.193 & 1.000 & 0.634 \\
-0.296 & 0.099 & 0.634 & 1.000  \end{pmatrix}.
  \label{color_cor:eqn}
 \end{equation}

% \begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=5.5in]{output11/sigma_corner.pdf} 
%   \caption{Posterior contours for the parameters $\sigma$, the square root of the diagonal elements of $C_c$.
%   \label{sigma:fig}}
%\end{figure}

Each supernova is described by its parameters $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)=(\gamma^0_B-\gamma^0_V)k$, and
$E_{\gamma^1}(B-V)=(\gamma^1_B-\gamma^1_V)k_1$, as well as its grey offset
$\Delta$: their distributions for all Monte Carlo links for all supernovae are shown in Figure~\ref{perobject:fig}.
There is a core concentration in the  parameter-space, with around ten objects that occupy its outskirts.
Many outliers appear in the red tail of $E_{\gamma^0}(B-V)$, as would be expected for the infrequent selection of supernovae
heavily extinguished by host-galaxy dust.
Outliers  are also clearly distinguished in  $EW_{Ca}$--$\lambda_{Si}$ space.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output11/perobject_corner.pdf} 
   \caption{Distributions for the supernova parameters $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, and $E_{\gamma^1}(B-V)$, as well as the grey offset
$\Delta$.  All links are plotted, so that each supernova contributes a cloud of points.
   \label{perobject:fig}}
\end{figure}

The Pearson correlation coefficients for $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_\gamma(B-V)$, and $E_\delta(B-V)$ are given in the matrix
\begin{equation}
\begin{pmatrix}
1.000 & 0.001 & -0.059 & -0.045 & 0.105 & 0.007 \cr
0.001 & 1.000 & 0.115 & -0.252 & -0.066 & -0.003 \cr
-0.059 & 0.115 & 1.000 & -0.140 & -0.163 & -0.024 \cr
-0.045 & -0.252 & -0.140 & 1.000 & 0.040 & 0.027 \cr
0.105 & -0.066 & -0.163 & 0.040 & 1.000 & 0.055 \cr
0.007 & -0.003 & -0.024 & 0.027 & 0.055 & 1.000
\end{pmatrix}.
\end{equation}
\color{red}
We find weak correlations between our color-excess parameters and the input features, the strongest being $-0.163$ between
$E_\gamma(B-V)$ and $EW_{Si}$;
\color{black}
recall that $EW_{Si}$ is correlated with light-curve shape, which is correlated with host-galaxy (including dust) properties 
\citep{2003MNRAS.340.1057S}.
An extensive discussion of the correlations between the spectral features of the SNfactory data set can be found in \citet{chotard:thesis}
and \citet{leget:thesis}.

\subsection{Two Extrinsic Parameters}
\color{red}
The two sets of  $\gamma^i$ parameters  are significantly non-zero. 
\color{black}
None of the 20000 links of 
our Monte Carlo chains for $\delta$ exceed 0 (see Figures~\ref{global1:fig}--\ref{global5:fig}), we thus claim detection of the
influence of $k$ and $k_1$  on supernova magnitudes
with probability $(1-5\times 10^{-5})$.

\color{red}
Combining the effects of both parameters, we can define a total $X$-band extinction $A_X=\gamma^0_X k + \gamma^1_X k_1$
and a corresponding $E(B-V) = A_B-A_V$.  Figure~\ref{avebv:fig} shows the relationship between $A_V$ and $E(B-V)$, from
which a linear fit gives a slope of $R_V=2.51^{+0.23}_{-0.21}$.  The fit also gives the zeropoint, which we use to 
estimate $R_V=A_V/(E(B-V)-zeropoint$ per supernova.  These values are shown in Figure~\ref{avebv:fig}.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.5in]{output11/avebv.pdf}
   \includegraphics[width=2.5in]{output11/rv.pdf}
   \caption{\textcolor{red}{Left: $A_V$ versus $E(B-V)$ and Right: $R_V=A_V/(E(B-V)-zeropoint$\
   versus $E(B-V)$ for the supernova sample.}
   \label{avebv:fig}}
\end{figure}

Until now we have considered $k$ and $D$ as independent external and internal parameters.  However, they may be both
parameters of a single underlying dust.  Indeed the CCM dust law can be exactly written as a two-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V).
\end{equation}
The Fitzpatrick model can also approximately be described by a linear model in the range of $R_V$ and wavelength under consideration:
the results shown here vary only at the percent level when making this approximation.
Similarly our model can be expressed as
\begin{equation}
A_X = \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2}  E_{\gamma^0}(B-V) +  \frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2}  E_{\gamma^1}(B-V).
\end{equation}

%
%
%The extrinsic parameters $\gamma$ attributed to dust are fit for without reference to a specific dust model:
%we now do so in a fit to the 
%model of \citet{1999PASP..111...63F},
%taking the fixed 3693,  4369,  5287,  6319, 7611 \AA\  for the effective wavelengths of the supernova flux through the bands.
%Figure~\ref{fitz:fig} shows the   68\% credible intervals of $\gamma$ and our result
%$R_V=2.97^{+0.18}_{-0.17}$..  
%We do the analysis in $\gamma$, as opposed to $R_V$ say, because the posteriors of $\gamma$ are well-approximated by a normal distribution that is used for the fit.
%The posterior distributions for $R_{\gamma X}=\frac{\gamma_X}{\gamma_B-\gamma_V}$ are shown in Figure~\ref{rx:fig}.
%\color{black}
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=3.2in]{output11/fitz.pdf}
%   \caption{\textcolor{red}{68\% credible intervals of $\gamma_X$ and the best-fit to the  \citet{1999PASP..111...63F}  model $R^F_V=2.97$.}
%   \label{fitz:fig}}
%\end{figure}
%
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=2.8in]{output11/rx_corner.pdf}
%      \includegraphics[width=2.8in]{output11/rxdelta_corner.pdf} 
%   \caption{Posterior contours for Left:  $R_X=\frac{\gamma_X}{\gamma_B-\gamma_V}$;  Right:  $R_{\delta X}=\frac{\delta_X}{\delta_B-\delta_V}$.
%   \label{rx:fig}}
%\end{figure}

@@

\textcolor{red}{
The values of $\delta_X/\delta_V$, plotted in  Figure~\ref{deltaratio:fig}, are not monotonic with color: they increase from the $U$ to $V$, and then decrease
from $V$ to $I$,
a consequence
of which is that
a reddening in $B-V$ comes with a blueing in $V-R$.
Also apparent is that a reddening of $B-V$ is associated with a brightening of the supernova.
The posterior distributions for $R_{\delta X}=\frac{\delta_X}{\delta_B-\delta_V}$ are shown in Figure~\ref{rx:fig}.
Although the posteriors of $\delta$ are close to Gaussian, $\delta_B-\delta_V$ can approach zero, resulting in the
elongated tail in negative  $R_{\delta X}$.
All values are significantly negative.}
These are not the behaviors expected from normal dust,
circumstellar dust suggested as a source of SN extinction \citep{2005ApJ...635L..33W,2008ApJ...686L.103G,
2015ApJ...807L..26G}, or a blackbody in the Raleigh-Jeans tail, and so may be indicative of
more complicated underlying supernova physics.

@@

@@
